---
title: "2wk vanc correlations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("phyloseq")
library("Biostrings")
library("ggplot2")
library("readxl")
library("tidyverse")
library("reshape2")
library("gridExtra")
library("ggrepel")
library("ggpubr")

setwd("~/code/Stool")

theme_set(theme_bw()) # optional, adjusts ggplot2 theme
```


```{r, include=FALSE}
# read in saved files
ps.2wkvanc <- readRDS("~/code/Stool/ps_2wkvanc.RDS")

```



# Just 2wk dataset
```{r}
# "Actinobacteria", "Bacteroidetes", "Firmicutes", "Proteobacteria", "Verrucomicrobia"
actino.2wk <- subset_taxa(ps.2wkvanc, Phylum=="Actinobacteria")
bact.2wk <- subset_taxa(ps.2wkvanc, Phylum=="Bacteroidetes")
firm.2wk <- subset_taxa(ps.2wkvanc, Phylum=="Firmicutes")
proteo.2wk <- subset_taxa(ps.2wkvanc, Phylum=="Proteobacteria")
verru.2wk <- subset_taxa(ps.2wkvanc, Phylum=="Verrucomicrobia")

```


```{r}
sum.2wkvanc <- estimate_richness(ps.2wkvanc, measures = c("Choa1", "Shannon", "Simpson"))
actino.sums <- as(sample_sums(actino.2wk),"matrix")
bact.sums <- as(sample_sums(bact.2wk),"matrix")
firm.sums <- as(sample_sums(firm.2wk),"matrix")
proteo.sums <- as(sample_sums(proteo.2wk),"matrix")
verru.sums <- as(sample_sums(verru.2wk),"matrix")
taxasums <- as(sample_sums(ps.2wkvanc), "matrix")

allsums <- cbind(actino.sums, bact.sums, firm.sums, proteo.sums, verru.sums, taxasums)
allsums <- as.data.frame(allsums)
names(allsums) <- c("Actinobacteria", "Bacteroidetes", "Firmicutes", "Proteobacteria", "Verrucomicrobia", "All")
allsums
```

# relative
create alternative see if relative valuse change anything
```{r}
taxasums <- as(sample_sums(ps.2wkvanc), "matrix")
relsums <- allsums/taxasums
relsums
```





```{r}
div.2wk <- estimate_richness(ps.2wkvanc, measures = c("Chao1", "Shannon", "Simpson"))
div.2wk

pdata.2wk <- sample_data(ps.2wkvanc)
pdata.2wk
```
```{r}
# combine pdata with richness/diversity data
pdiv.2wk <- cbind(pdata.2wk, div.2wk)
# clear row names (row names are auto set to sample id, but there's a column for that)
row.names(pdiv.2wk) <- NULL
# drop unnessary columns
pdiv.2wk <- select(pdiv.2wk, -Experiment, -Condition, -Genotype, -Collection.Date, -Sex)
pdiv.2wk
```

```{r}
csv_out <- cbind(pdiv.2wk, allsums)
rel_out <- cbind(pdiv.2wk, relsums)

row.names(csv_out) <- NULL
row.names(rel_out) <- NULL

csv_out
rel_out

write.csv(csv_out, "~/code/Stool/csv/2wk_corr_start.csv", quote=FALSE)
write.csv(rel_out, "~/code/Stool//csv/2wk_rel_start.csv", quote=FALSE)

```


*******************************************************
# START HERE!!!
*******************************************************

```{r}
# Read in delta data of div and bacteria counts
deltas <- read.csv("~/code/Stool/csv/2wk_corr_deltas.csv")
reldeltas <- read.csv("~/code/Stool/csv/2wk_rel_deltas.csv")


# read in pheno data of mice (includes raw values and deltas)
pheno.2wk <- read.csv("~/code/Stool/csv/2wk_pheno.csv")


# and clean combine pheno data with other delta datas
pheno.2wk <- select(pheno.2wk, -MMA.baseline, -MMA.2wk, -Fgf21.baseline, -Fgf21.2wk, -Weight.baseline, -Weight.2wk)
deltas <- select(deltas, -se.chao1)
reldeltas <- select(reldeltas, -se.chao1)

all_data <- left_join(pheno.2wk, deltas, by="Mouse.ID")
all_data_rel <- left_join(pheno.2wk, reldeltas, by="Mouse.ID")
all_data
all_data_rel
```

```{r}
# all_data <- select(all_data, -MMA.baseline, -MMA.2wk, -Fgf21.baseline, -Fgf21.2wk, -Weight.baseline, -Weight.2wk, -se.chao1)
names(all_data)
```

# correlation metrix - raw abundances
```{r}
# correlation matrix
pairs(all_data)

png(filename="~/code/Stool/pics/2wk_correlation_matrix.png", height=10, width=10, units="in", res=300)
pairs(all_data)
dev.off()
```

# correlation matrix - relative abundances
```{r}
# correlation matrix
pairs(all_data_rel)

png(filename="~/code/Stool/pics/2wk_rel_correlation_matrix.png", height=10, width=10, units="in", res=300)
pairs(all_data_rel)
dev.off()
```





```{r}
b <- ggplot(deltas, aes(x=Shannon, y=Bacteroidetes)) +
  geom_point() +
  geom_label(label=deltas$Mouse.ID, size=3, nudge_x = -0.2, nudge_y = 750)
  
b
```


```{r}
p <- ggplot(deltas, aes(x=Shannon, y=Proteobacteria)) +
  geom_point(size=5) +
  # geom_label(label=deltas$Mouse.ID, size=3, nudge_x = -0.2, nudge_y = 750)
  geom_text_repel(label=deltas$Mouse.ID, size=5)+
  geom_smooth(method=lm, linetype="dashed", color="darkred") #+
  # stat_regline_equation
  
p
```

ggscatter(df, x = "wt", y = "mpg",
   color = "black", shape = 21, size = 3, # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")
   )

```{r}
lnreg <- ggscatter(all_data, x=micros[1], y=phenos[1], 
                   size = 5, color= colors[1],
                  add = "reg.line",  # Add regressin line
                  add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
                   conf.int = FALSE, # Add confidence interval
                   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
                   cor.coeff.args = list(method = "pearson", label.sep = "\n"),
                  cor.coef.size = 5,
                  
                  ) +
     ggtitle(paste0(phenos[1], " x ", micros[1])) +
     theme(plot.title = element_text(hjust = 0.5, size=24))+
     geom_text_repel(label=all_data$Experiment.Details, size=3)

lnreg
```






```{r}
br <- ggplot(reldeltas, aes(x=Shannon, y=Bacteroidetes)) +
  geom_point()# +
  # geom_label(label=reldeltas$Mouse.ID, size=3, nudge_x = -0.2, nudge_y = 750)
  
br
```


```{r}
pr <- ggplot(reldeltas, aes(x=Shannon, y=Proteobacteria)) +
  geom_point()# +
  # geom_label(label=reldeltas$Mouse.ID, size=3, nudge_x = -0.2, nudge_y = 750)
  
pr
```





**************************************************************************************
# Automate graph making process
**************************************************************************************

```{r}
phenos <- c("MMA.delta", "Fgf21.delta", "Weight.delta")
micros <- c("Shannon", "Simpson", "Chao1", "Actinobacteria", "Bacteroidetes", "Firmicutes", "Proteobacteria", "Verrucomicrobia", "All")
colors <- c("black", "dimgrey", "grey", "#F8766D", "#C49A01", "#53B400", "#01B6EB", "#FB60D7", "brown"
)
```


# create correlation graphs
```{r}
for (i in seq_along(phenos)){
  for (j in seq_along(micros)){
   plot <- ggplot(all_data, aes_string(x=micros[j], y=phenos[i])) +
     geom_point(color=colors[j], size=5) +
     geom_text_repel(label=all_data$Mouse.ID, size=5) +
     ggtitle(paste0(phenos[i], " x ", micros[j])) +
     theme(plot.title = element_text(hjust = 0.5, size=24))
   
   ggsave(plot, file=paste0("~/code/Stool/pics/correl_pics/", phenos[i], "_", micros[j], ".png"))
  }
}
```


for (i in seq_along(gois)) {

  plot <- ggplot(filter(labeled_gois_FPKM, gene == gois[i]), aes(x=group, y = FPKM, fill=group)) +
    stat_summary(geom = "bar", fun.y = mean, position = "dodge") +
    stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge", width = 0.2)  +
    stat_compare_means()+
    stat_compare_means(comparisons = my_comparisons, label = "p.signif")+
    theme_classic()+
    scale_fill_manual(values=colors) +
    ggtitle(gois[i]) +
    theme(plot.title = element_text(hjust = 0.5, size=24))

  ggsave(plot, file=paste("/Users/iskoec/code/qPCR_FPKM/pics/", gois[i], "_FPKM.tiff", sep=""))
  
  }
