---
title: "Dada2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dada2")
library("ggplot2")
library("tidyr")
library("dplyr")
library("stringr")
setwd("~/code/Stool")
```

## Dada2 Pipeline

See [dada2 github](https://benjjneb.github.io/dada2/index.html) for more details including tutorial (used to construct this pipeline).

Note, loaded dada2 from anaconda cloud, v1.14.0

Assumptions made in making this pipeline:
* samples were already demultiplexed, (true, cuz reads split into individual per-sample fastq files)
* adapters were removed, 
* and sequences paired (??? only one fastq files, no read 1/2 aka read f/r)
prior to sending fastq sequences

### Define paths

```{r}
# create list of sample locations
path <- "/Users/iskoec/code/Stool/MS0051-MMAmouse"
samples <- sort(list.files(path, pattern = ".fastq", full.names = TRUE))
samples

```


```{r}
# create list of sample names
sample.names <- sapply(strsplit(basename(samples), ".fastq"), `[`, 1)
sample.names
```



### Inspect read quality profiles
```{r}
plotQualityProfile(samples[1:2])
```


### Filter and trim
***note:*** had trouble running learn/plot errors bc had non A/C/T/G in seqs


```{r}
# Place filtered files in filtered/ subdirectory
filtSamples <- file.path(path, "filtered", paste0(sample.names, "filt.fastq.gz"))
filtSamples
names(filtSamples) <- sample.names
```


```{r}
out <- filterAndTrim(samples, filtSamples, maxN=0, compress = TRUE)
head(out)
```




### Learn error rates
```{r}
# takes about 10 mins to run
err <- learnErrors(filtSamples)
plotErrors(err, nominalQ = TRUE)
```

### save progress, so don't have to rerun
```{r}
saveRDS(filtSamples, "~/code/Stool/filtSamples.rds")
saveRDS(err, "~/code/Stool/err.rds")
```





## START HERE - DON'T HAVE TO REGENERATE/RECALCULATE FILTERING AND ERROR
```{r}
filtSamples <- readRDS("~/code/Stool/filtSamples.rds")
err <- readRDS("~/code/Stool/err.rds")
plotErrors(err, nominalQ = TRUE)
```






