---
title: "V13 cleaned all phyloseq"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("phyloseq")
library("Biostrings")
library("ggplot2")
library("readxl")
library("tidyverse")

setwd("~/code/Stool")

theme_set(theme_bw()) # optional, adjusts ggplot2 theme
```

Following last part of dada2 tutorial see:
https://benjjneb.github.io/dada2/tutorial.html

create dataframe with pdata
```{r}
# samdf = pData for samples
samdf <- read.csv("~/code/Stool/MS0067_MMAMicro/pData_MMAMicro.csv")
seqtab.nochim <- readRDS("~/code/Stool/MS0067_MMAMicro/mouse/seqtab.nochim.RDS")
taxa <- readRDS("~/code/Stool/MS0067_MMAMicro/mouse/taxa.RDS")

# Match sample names across tables
samples.out <- rownames(seqtab.nochim)
rownames(samdf) <- samples.out

ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows = FALSE),
               sample_data(samdf),
               tax_table(taxa))

```


Rename taxa to a short string and save ASVs in refseq slot
```{r}
dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))
ps
```

# Write out OTU table
```{r}
ps.otu <- as(otu_table(ps), "matrix")
ps.otu <- as_data_frame(ps.otu)
div <- estimate_richness(ps, measures="Shannon")
csv_out <- cbind(div, ps.otu)
write.csv(csv_out, "~/code/Stool/div_otu.csv")
```



```{r}
# exclude sample that does not have reads
ps <- prune_samples(sample_names(ps)!="VXV16421", ps)
# exclude control and fish data
ps <- subset_samples(ps, Experiment%in%c("2_wk_vancomycin", "4_week_Cocktail", "Vancomycin_Redose"))
ps.pdata <- data.frame(sample_data(ps))
ps.rel <- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))

# restrict phyla so graphs easier to read
ps.phyla <- subset_taxa(ps, Phylum%in%c("Actinobacteria", "Bacteroidetes", "Firmicutes", "Proteobacteria", "Verrucomicrobia", "Fusobacteria", "Tenericutes"))
ps.phyla.rel <- subset_taxa(ps.rel, Phylum%in%c("Actinobacteria", "Bacteroidetes", "Firmicutes", "Proteobacteria", "Verrucomicrobia", "Fusobacteria", "Tenericutes"))
```

# alpha diversity - all experiments
```{r}
alpha <- plot_richness(ps, x = "Experiment", measures=c("Chao1","Shannon", "Simpson"), color="Experiment.Details")
alpha
```
```{r}
ggsave("all_alpha.png", plot=alpha, path="~/code/Stool/pics", width=10, height=7)
```



# Beta diveristy - all experiments
```{r}
# transform data to proportions (relative abundances) as appropriate for Bray-Curtis distances 
ps.ord <- subset_samples(ps, Experiment != "Control")
ps.prop <- transform_sample_counts(ps.ord, function(otu) otu/sum(otu))
ord.nmds.bray <- ordinate(ps.prop, method="NMDS", distance="bray")
p_ordin<- plot_ordination(ps.prop, ord.nmds.bray, color="Experiment.Details", title="Bray NMDS") + facet_wrap(~Experiment, scale="free_x")
p_ordin
```
```{r}
ggsave("all_beta.png", plot=p_ordin, path="~/code/Stool/pics", width=10, height=6)
```



# absolute abundance - all experiments
```{r}
ps.bar <- plot_bar(ps.phyla, x="Sample.ID", fill="Phylum") + facet_wrap(~Experiment, scale="free_x")
```
```{r}
ps.bar <- ps.bar + labs(title="All experiments - absolute", subtitle= "Absolute abundances", x="Mouse ID") + 
  scale_x_discrete(breaks=ps.pdata$Sample.ID, labels=ps.pdata$Mouse.ID)
ps.bar
```
```{r}
ggsave("all_abund.png", plot = ps.bar, path="~/code/Stool/pics", width=10, height=6)
```



# relative abundance - all experiments
```{r}
ps.bar.rel <- plot_bar(ps.phyla.rel,  x="Sample.ID", fill="Phylum") + facet_wrap(~Experiment, scale="free_x")
```
```{r}
ps.bar.rel <-ps.bar.rel + labs(title="All experiments", subtitle = "Relative Abundances", x="Mouse ID") + 
  scale_x_discrete(breaks=ps.pdata$Sample.ID, labels=ps.pdata$Mouse.ID)
ps.bar.rel
```
```{r}
ggsave("all_abund_rel.png", plot = ps.bar, path="~/code/Stool/pics", width=10, height=6)

```




# 4 wk cocktail
```{r}
ps.4wk <- subset_samples(ps.phyla, Experiment=="4_week_Cocktail")
ps.4wk.pdata <- data.frame(sample_data(ps.4wk))
ps.4wk.rel <- subset_samples(ps.phyla.rel, Experiment=="4_week_Cocktail")

ps.4wk.bar <- plot_bar(ps.4wk, x="Sample.ID", fill="Phylum")
ps.4wk.rel.bar <- plot_bar(ps.4wk.rel, x="Sample.ID", fill="Phylum")
```
## absolute abundance
```{r}
ps.4wk.bar <- ps.4wk.bar + facet_wrap(~Condition, scale="free_x") +
  labs(title="4 Week Antibiotic Cocktail", subtitle="End of first treatment week", x="Mouse ID") + 
  scale_x_discrete(breaks=ps.4wk.pdata$Sample.ID, labels=ps.4wk.pdata$Mouse.ID)
ps.4wk.bar
```
```{r}
ggsave("4wk_abund.png", plot = ps.4wk.bar, path="~/code/Stool/pics", width=10, height=7)
```

## relative abundance
```{r}
ps.4wk.rel.bar <- ps.4wk.rel.bar + facet_wrap(~Condition, scale="free_x") +
  labs(title="4 Week Antibiotic Cocktail", subtitle="End of first treatment week", x="Mouse ID") + 
  scale_x_discrete(breaks=ps.4wk.pdata$Sample.ID, labels=ps.4wk.pdata$Mouse.ID)
ps.4wk.rel.bar
```
```{r}
ggsave("4wk_abund_rel.png", plot = ps.4wk.rel.bar, path="~/code/Stool/pics", width=10, height=7)

```




# Redose vancomycin
```{r}
ps.redose <- subset_samples(ps.phyla, Experiment=="4_week_Cocktail")
ps.redose.pdata <- data.frame(sample_data(ps.redose))
ps.redose.rel <- subset_samples(ps.phyla.rel, Experiment=="4_week_Cocktail")

ps.redose.bar <- plot_bar(ps.redose, x="Sample.ID", fill="Phylum")
ps.redose.rel.bar <- plot_bar(ps.redose.rel, x="Sample.ID", fill="Phylum")
```
## absolute abundance
```{r}
ps.redose.bar <- ps.redose.bar + facet_wrap(~Condition, scale="free_x") +
  labs(title="4 Week Antibiotic Cocktail", subtitle="End of first treatment week", x="Mouse ID") + 
  scale_x_discrete(breaks=ps.redose.pdata$Sample.ID, labels=ps.redose.pdata$Mouse.ID)
ps.redose.bar
```
```{r}
ggsave("redose_abund.png", plot = ps.redose.bar, path="~/code/Stool/pics", width=10, height=7)

```


## relative abundance
```{r}
ps.redose.rel.bar <- ps.redose.rel.bar + facet_wrap(~Condition, scale="free_x") +
  labs(title="4 Week Antibiotic Cocktail", subtitle="End of first treatment week", x="Mouse ID") + 
  scale_x_discrete(breaks=ps.redose.pdata$Sample.ID, labels=ps.redose.pdata$Mouse.ID)
ps.redose.rel.bar
```
```{r}
ggsave("redose_abund_rel.png", plot = ps.redose.rel.bar, path="~/code/Stool/pics", width=10, height=7)

```





