---
title: "V13 2wk vanc"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("phyloseq")
library("Biostrings")
library("ggplot2")
library("readxl")
library("tidyverse")
library("reshape2")
library("gridExtra")

setwd("~/code/Stool")

theme_set(theme_bw()) # optional, adjusts ggplot2 theme
```

Following last part of dada2 tutorial see:
https://benjjneb.github.io/dada2/tutorial.html

Read in data from dada2 pipeline
```{r}
# samdf = pData for samples
samdf <- read.csv("~/code/Stool/MS0067_MMAMicro/pData_MMAMicro.csv")
seqtab.nochim <- readRDS("~/code/Stool/MS0067_MMAMicro/mouse/seqtab.nochim.RDS")
taxa <- readRDS("~/code/Stool/MS0067_MMAMicro/mouse/taxa.RDS")

# Match sample names across tables
# note: rownames must be sample names for all tables for phyloseq!***
samples.out <- rownames(seqtab.nochim)
rownames(samdf) <- samples.out


ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows = FALSE),
               sample_data(samdf),
               tax_table(taxa))
```

*** Seperate phylo object into experiments? ***


Rename taxa to a short string and save ASVs in refseq slot
```{r}
dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))
ps

# # exclude sample that does not have reads
# ps <- prune_samples(sample_names(ps)!="VXV16421", ps)
# ps
```

# Subset data to 2 wk vancomycin experiment
```{r}
ps.2wkvanc <- subset_samples(ps, Experiment == "2_wk_vancomycin")
ps.2wk.pData <- sample_data(ps.2wkvanc)

```


# Write out otu table for 2wk vanc
```{r}
otu.2wk <- as.data.frame(as(otu_table(ps.2wkvanc), "matrix"))
div.2wk <- estimate_richness(ps.2wkvanc, measure="Shannon")
csv_2wk <- cbind(div.2wk,otu.2wk)
write.csv(csv_2wk, file="~/code/Stool/2wk_div_otu.csv", quote=FALSE)

```



# Visualize alpha-diversity
```{r}
p_alpha <- plot_richness(ps.2wkvanc, x = "Experiment.Details", measures=c("Chao1", "Shannon", "Simpson"), color="Experiment.Details", shape="Sex") +
  geom_line(aes(group = Mouse.ID), colour = "black") +
  labs(x="Experiment Timepoint")
p_alpha
```
# write out png
```{r}
ggsave("2wkvanc_alpha.png", plot=p_alpha, path="~/code/Stool/pics", height = 4, width=6)
```


```{r}
p_alpha_labs2 <- p_alpha + geom_label(mapping = aes(label=Mouse.ID), size=1)
p_alpha_labs2
```


```{r}
p_alpha_labs <- plot_richness(ps.2wkvanc, x = "Experiment.Details", measures=c( "Shannon"), color="Experiment.Details", shape="Sex") +
  geom_line(aes(group = Mouse.ID), colour = "black") +
  labs(x="Experiment Timepoint") + 
  geom_label(label="2982/7368 F1 -/-", x=1.75, y=2.15, size=2, color="black", fill="#00BA38")
p_alpha_labs
```
# write out png
```{r}
ggsave("2wkvanc_alpha_lab.png", plot=p_alpha_labs, path="~/code/Stool/pics", height = 4, width=3.55)
```








# check diversity of baseline values
```{r}
ps.2wk.baseline <- subset_samples(ps.2wkvanc, Experiment.Details!="Treatment_wk2")
palpha_base <- plot_richness(ps.2wk.baseline, x = "Experiment.Details", measures="Shannon", color="Experiment.Details", shape="Sex", title="Shannon Index") +
  facet_wrap(~Mouse.ID)
palpha_base
```





# output alpha diversity
```{r}
adiv <- estimate_richness(ps.2wkvanc, measures = c("Shannon", "Simpson"))
adiv
```
```{r}
ps.prop <- transform_sample_counts(ps.2wkvanc, function(otu) otu/sum(otu))
ord.nmds.bray <- ordinate(ps.prop, method="NMDS", distance="bray")
```
```{r}
p_ordin<- plot_ordination(ps.prop, ord.nmds.bray, color="Experiment.Details", title="Bray NMDS", shape = "Sex")
p_ordin
```
# save pic
```{r}
ggsave("2wkvanc_beta.png", plot = p_ordin, path = "~/code/Stool/pics", width=6, height=4)
```


# label outlier
```{r}
p_ordin_lab <- plot_ordination(ps.prop, ord.nmds.bray, color="Experiment.Details", title="Bray NMDS", shape = "Sex") +
  geom_text(mapping = aes(label=Mouse.ID), size = 2.5, hjust = 0, nudge_x = -1000)
p_ordin_lab
```

```{r}
p_ordin_outlier <- plot_ordination(ps.prop, ord.nmds.bray, color="Experiment.Details", title="Bray NMDS", shape = "Sex")+
  geom_label(label="2982/7368 F1 -/-", x=1800, y=0.085, size=2, color="black", fill="#00BA38")
p_ordin_outlier
```

# save pic
```{r}
ggsave("2wkvanc_beta_outlier.png", plot = p_ordin_outlier, path = "~/code/Stool/pics", width=6, height=4)
```










# Absolute abundances
```{r}
abs <- plot_bar(ps.2wkvanc, x="Sample.ID", fill="Phylum")
```
# show graph
```{r}
abs <- abs + facet_wrap(~Experiment.Details, scale="free_x") +
  scale_x_discrete(breaks=ps.2wk.pData$Sample.ID, labels=ps.2wk.pData$Mouse.ID)+
  labs(x="Mouse ID")
abs
```
# save pic
```{r}
ggsave(file="2wkvanc_abs_abund.png", plot = abs, path="~/code/Stool/pics", width = 10, height=7)
```





# relative abundances
```{r}
ps.2wk.rel <- transform_sample_counts(ps.2wkvanc, function(OTU) OTU/sum(OTU))
rel <- plot_bar(ps.2wk.rel, x="Sample.ID", fill="Phylum")
```
# relative abundance graph adjustments
```{r}
rel <- rel + facet_wrap(~Experiment.Details, scale="free_x") +
  scale_x_discrete(breaks=ps.2wk.pData$Sample.ID, labels=ps.2wk.pData$Mouse.ID)
rel
```

# Restrict output to 5 phyla
See following link for script function to easily extract abundances from phyloseq object:
https://rdrr.io/github/microbiome/microbiome/src/R/abundances.R
```{r}
ps.phyla <- subset_taxa(ps.2wkvanc, Phylum%in%c("Actinobacteria", "Bacteroidetes", "Firmicutes", "Proteobacteria", "Verrucomicrobia"))

startSums <- data.frame(sample_sums(ps.2wkvanc))
endSums <- data.frame(sample_sums(ps.phyla))
start_end <- cbind(startSums, endSums)

start_end <- mutate(start_end, percent.change = (sample_sums.ps.phyla.-sample_sums.ps.2wkvanc.)/sample_sums.ps.phyla.*100)
start_end

hist(start_end$percent.change)

df_print <- cbind(ps.2wk.pData$Mouse.ID, ps.2wk.pData$Experiment.Details, start_end)
names(df_print) <- c("Mouse ID", "Treatment Timepoint", "Raw Sample Sums", "Pruned Sample Sums", "Percent Change")
df_print
```
# save df for ppt
```{r}
png(filename = "~/code/Stool/pics/pruned_nums.png", width=9.5, height=7.5, units="in", res=300)
grid.table(df_print)
dev.off()
```


# Absolute abundances
```{r}
abs_phyla <- plot_bar(ps.phyla, x="Sample.ID", fill="Phylum")
```
# show graph
```{r}
abs_phyla <- abs_phyla + facet_wrap(~Experiment.Details, scale="free_x") +
  scale_x_discrete(breaks=ps.2wk.pData$Sample.ID, labels=ps.2wk.pData$Mouse.ID)+
  labs(x="Mouse ID")
abs_phyla
```
```{r}
ggsave("2wkvanc_abs_phyla_abund.png", plot=abs_phyla, path = "~/code/Stool/pics", width = 10, height=7)
```




# relative abundances
```{r}
ps.2wk.rel <- transform_sample_counts(ps.phyla, function(OTU) OTU/sum(OTU))
rel_phyla <- plot_bar(ps.2wk.rel, x="Sample.ID", fill="Phylum")
```
# relative abundance graph adjustments
```{r}
rel_phyla <- rel_phyla + facet_wrap(~Experiment.Details, scale="free_x") +
  scale_x_discrete(breaks=ps.2wk.pData$Sample.ID, labels=ps.2wk.pData$Mouse.ID)+
  labs(x="Mouse ID")
rel_phyla
```

```{r}
ggsave("2wkvanc_rel_phyla_abund.png", plot=rel_phyla, path = "~/code/Stool/pics", width = 10, height=7)
```



