---
title: "V13 phyloseq"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("phyloseq")
library("Biostrings")
library("ggplot2")
library("readxl")
library("tidyverse")

setwd("~/code/Stool")

theme_set(theme_bw()) # optional, adjusts ggplot2 theme
```

Following last part of dada2 tutorial see:
https://benjjneb.github.io/dada2/tutorial.html

create dataframe with pdata
```{r}
# samdf = pData for samples
samdf <- read.csv("~/code/Stool/MS0067_MMAMicro/pData_MMAMicro.csv")
seqtab.nochim <- readRDS("~/code/Stool/MS0067_MMAMicro/mouse/seqtab.nochim.RDS")
taxa <- readRDS("~/code/Stool/MS0067_MMAMicro/mouse/taxa.RDS")
```

Match sample names across tables
*** note: rownames must be sample names for all tables for phyloseq!***
```{r}
samples.out <- rownames(seqtab.nochim)
rownames(samdf) <- samples.out
```



```{r}
ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows = FALSE),
               sample_data(samdf),
               tax_table(taxa))
```

*** Seperate phylo object into experiments? ***


Rename taxa to a short string and save ASVs in refseq slot
```{r}
dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))
ps

# exclude sample that does not have reads
ps <- prune_samples(sample_names(ps)!="VXV16421", ps)
ps
```



# Visualize alpha-diversity
```{r}
plot_richness(ps, x = "Experiment", measures=c("Chao1","Shannon", "Simpson"), color="Experiment.Details")
```

```{r}

```



# Ordinate -visualize beta diversity
- had trouble here, possibly due to empty rows from water controls, removed controls: https://github.com/joey711/phyloseq/issues/851

```{r}
# transform data to proportions (relative abundances) as appropriate for Bray-Curtis distances 
ps.ord <- subset_samples(ps, Experiment != "Control")
ps.prop <- transform_sample_counts(ps.ord, function(otu) otu/sum(otu))
ord.nmds.bray <- ordinate(ps.prop, method="NMDS", distance="bray")
```

```{r}
p_ordin<- plot_ordination(ps.prop, ord.nmds.bray, color="Experiment.Details", title="Bray NMDS") + facet_wrap(~Experiment, scale="free_x")
p_ordin
```



```{r}

top20 <- names(sort(taxa_sums(ps), decreasing=TRUE))[1:20]
ps.top20 <- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
ps.top20 <- prune_taxa(top20, ps.top20)
plot_bar(ps.top20, x="Sample.ID", fill="Phylum") + facet_wrap(~Experiment, scale="free_x")
```

# Raw abundances
```{r}
ps.ab <- subset_samples(ps, Experiment%in%c("2_wk_vancomycin", "4_week_Cocktail", "Vancomycin_Redose"))
plot_bar(ps.ab, x="Sample.ID", fill="Phylum") + facet_wrap(~Experiment, scale="free_x")
```

# Relative abundances
```{r}
ps.ab.rel <- transform_sample_counts(ps.ab, function(OTU) OTU/sum(OTU))
plot_bar(ps.ab.rel, x="Sample.ID", fill="Phylum") + facet_wrap(~Experiment, scale="free_x")
```

# Subset taxa to those seen in previous experiment, and calculate relative abundances

```{r}
ps.rel <- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
ps.phyla <- subset_taxa(ps.rel, Phylum%in%c("Actinobacteria", "Bacteroidetes", "Firmicutes", "Proteobacteria", "Verrucomicrobia", "Fusobacteria", "Tenericutes"))

```



# 4 Week Cocktail alone
## absolute abundance
```{r}
ps.ab.4wk <- subset_samples(ps, Experiment=="4_week_Cocktail")
ps.ab.4wk.df <- data.frame(sample_data(ps.ab.4wk))
ps.phyla.4wk <- subset_taxa(ps.ab.4wk, Phylum%in%c("Actinobacteria", "Bacteroidetes", "Firmicutes", "Proteobacteria", "Verrucomicrobia", "Fusobacteria", "Tenericutes"))

p.4wk.bar <- plot_bar(ps.phyla.4wk, x="Sample.ID", fill="Phylum")
```
```{r}
p.4wk.bar <- p.4wk.bar + facet_wrap(~Condition, scale="free_x") +
  labs(title="4 Week Antibiotic Cocktail", subtitle="End of first treatment week", x="Mouse ID") + 
  scale_x_discrete(breaks=ps.ab.4wk.df$Sample.ID, labels=ps.ab.4wk.df$Mouse.ID)
p.4wk.bar
```
```{r}
ggsave("4wk_abs_abun.png", plot=p.4wk.bar, path = "~/code/Stool/pics", width=10, height=7)
```




## relative abundance
```{r}
ps.4wk <- subset_samples(ps.phyla, Experiment == "4_week_Cocktail")
ps.4wk.pData <- data.frame(sample_data(ps.4wk))
p_4wk <- plot_bar(ps.4wk, x="Sample.ID", fill="Phylum")

```

```{r}
p_4wk + 
  facet_wrap(~Condition, scale="free_x") +
  labs(title="4 Week Antibiotic Cocktail", subtitle="End of first treatment week", x="Mouse ID") + 
  scale_x_discrete(breaks=ps.4wk.pData$Sample.ID, labels=ps.4wk.pData$Mouse.ID)
  
```



# Vancomycin redose
```{r}
ps.revanc <- subset_samples(ps.phyla, Experiment == "Vancomycin_Redose")

p_revanc <- plot_bar(ps.revanc, x="Sample.ID", fill="Phylum")
```

```{r}
p_revanc + labs(title="Vancomycin_Redose") + facet_wrap(~Experiment.Details, scale="free_x") +
  scale_x_discrete(breaks=ps.ab.4wk.df$Sample.ID, labels=ps.ab.4wk.df$Mouse.ID)

```






2 Week Vancomycin
```{r}
ps.2wk <- subset_samples(ps.phyla, Experiment == "2_wk_vancomycin")
ps.2wk.pData <- data.frame(sample_data(ps.2wk))

p_2wk <- plot_bar(ps.2wk, x="Sample.ID", fill="Phylum")
p_2wk + labs(title="2 week vancomycin Experiment", subtitle="RNAseq samples") + facet_wrap(~Experiment.Details, scale="free_x")+
  scale_x_discrete(breaks=ps.2wk.pData$Sample.ID, labels=ps.2wk.pData$Mouse.ID)
p_2wk
```

```{r}
ps.ab.phyla <- subset_taxa(ps.ab, Phylum%in%c("Actinobacteria", "Bacteroidetes", "Firmicutes", "Proteobacteria", "Verrucomicrobia", "Fusobacteria", "Tenericutes"))


ps.2wk <- subset_samples(ps.ab.phyla, Experiment == "2_wk_vancomycin")
ps.2wk.pData <- data.frame(sample_data(ps.2wk))

p_2wk <- plot_bar(ps.2wk, x="Sample.ID", fill="Phylum")
p_2wk + labs(title="2 week vancomycin Experiment", subtitle="RNAseq samples") + facet_wrap(~Experiment.Details, scale="free_x")+
  scale_x_discrete(breaks=ps.2wk.pData$Sample.ID, labels=ps.2wk.pData$Mouse.ID)
p_2wk
```

